\propertitle

### 2019-06-14 maxNt_v5.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-06-14T14:06:05+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal population\\ from subpopulation data%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

@@

These two steps are even more straightforward if instead of power moments
we use \emph{normalized factorial moments}. The $m$th normalized factorial
moment is defined as the expectation
\begin{equation}
  \E\condt[\Bigg]{
    \binom{n \ya}{m}\,\binom{n}{m}^{-1}}{\yH}
  \equiv
  \sum_{\ya} %m!\,
  \binom{n \ya}{m}\,  \binom{n}{m}^{-1}\, \pf(\ya \| \yH),
\end{equation}
where $\ya\in\set{0, 1/n, 2/n, \dotsc, (n-1)/n, 1}$ is the normalized total
activity of a population of $n$ neurons, and $\pf(\ya \| \yH)$ is the
probability for activity $\ya$ given assumptions $\yH$. This moment can be
interpreted as the expectation of the number of distinct $m$-tuples of
simultaneously active neurons, normalized by the maximum possible number of
distinct $m$-tuples. For example, with $m=2$, if $na=4$ neurons are active
in a population of $n=5$, we have $\tbinom{4}{2}=6$ distinct pairs of
simultaneously active neurons, and if all five neurons were active we would
have a maximum of $\tbinom{5}{2}=10$ distinct pairs; the normalized number
of pairs is therefore $6/10$. Note that the first $m$ factorial moments
linearly determine the first $m$ moments and vice versa, so fixing the ones is
equivalent to fixing the others. Now assume that these $n$ neurons are a
sample from a full population of $N$ neurons, with possible normalized
activity $\yA \in\set{0,1/N,\dotsc,1}$. We have this extremely convenient
property: \emph{The first $n-1$ normalized factorial moments for the sample
  and for the full population are numerically identical}:
\begin{equation}
  \label{eq:equal_m_factorialmoments}
  \E\condt[\big]{\tbinom{n\ya}{m}\;\tbinom{n}{m}^{-1}}{\yH} =
  \E\condt[\big]{\tbinom{N\yA}{m}\;\tbinom{N}{m}^{-1}}{\yH},
  \qquad m \le n.
\end{equation}

We can hence apply the maximum-entropy method to obtain a distribution for
the full population by constraining its $m$th factorial moment to be equal
to the sample's recorded average of $\tbinom{n\ya}{m}\;\tbinom{n}{m}^{-1}$,
for as many $m$ as we please with $1\le m \le n$. If $\ya_{t}$ is the
normalized total activity of the sample during the time bin $t$, during a
recording with $T$ time bins, the constraint on the distribution
$\pf(\yA \| \yH)$ for the full population is
\begin{equation}
  \label{eq:final_constraint}
  \frac{1}{T}\sum_{t}\binom{n\ya_{t}}{m}\,\binom{n}{m}^{-1}
  =
   \sum_{\yA}
  \binom{N \yA}{m}\,  \binom{N}{m}^{-1}\, \pf(\yA \| \yH)
\end{equation}
for any $m$s we wish.

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-06-14T15:23:13+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal population\\ from subpopulation data%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

Let's introduce the context and notation for our problem.

We have recorded the activities of $n$ neurons for a given amount of time.
These activities are binned into $T$ time bins and binarized to $\set{0,1}$
values. Denote by $\ya_{t}$ the number of active neurons during time bin
$t$, divided by the total number of neurons $n$; we call this the
\emph{normalized total activity} or just \enquote{total activity} for
short. Obviously $\ya_{t} \in\set{0,1/n,\dotsc,(n-1)/n,1}$: if $\ya_{t}=0$,
all $n$ neurons were silent during time bin $t$; if $\ya_{t}=1$, all were
active.

\mynote{maybe move this paragraph to the intro?}Present-day technologies
allow the recording of neuronal activity from small brain regions
\mynote{be more specific}. The specific neurons that happen to be recorded
from a region are not and cannot yet be specifically chosen, however, owing
to several factors -- for example, limitations in the precision with which
electrodes are inserted, or in the targeting of neurons by viruses, and so
on. In fact, the set of recorded neurons may vary during very long
recordings or across experimental trials. We assume that there's a small
region, comprising a population of $N$ neurons, for which other sets of
neurons had equal chances of being \enquote{selected} as the sample of $n$
neurons that have actually been recorded. During the recording, the
normalized total activity of these $N$ neurons at time bin $t$ was
$\yA_{t}$; note that $\yA_{t} \in \set{0,1/N,\dotsc,(N-1)/N,1}$ and that
$N\yA_{t} \ge n\ya_{t}$ for obvious reasons.

Our questions concern general features of the total activity $\yA$ of the
full population; for example: what was its time-frequency distribution? How
much activity should we expect in a time bin during the recording, or in a
new time bin? The answers to these questions are probabilistic and come in
the form of distributions.

@@

These two steps are even more straightforward if instead of power moments
we use \emph{normalized factorial moments}. The $m$th normalized factorial
moment is defined as the expectation
\begin{equation}
  \E\condt[\Bigg]{
    \binom{n \ya}{m}\,\binom{n}{m}^{-1}}{\yH}
  \equiv
  \sum_{\ya} %m!\,
  \binom{n \ya}{m}\,  \binom{n}{m}^{-1}\, \pf(\ya \| \yH),
\end{equation}
where $\ya\in\set{0, 1/n, 2/n, \dotsc, (n-1)/n, 1}$ is the normalized total
activity of a population of $n$ neurons, and $\pf(\ya \| \yH)$ is the
probability for activity $\ya$ given assumptions $\yH$. This moment can be
interpreted as the expectation of the number of distinct $m$-tuples of
simultaneously active neurons, normalized by the maximum possible number of
distinct $m$-tuples. For example, with $m=2$, if $na=4$ neurons are active
in a population of $n=5$, we have $\tbinom{4}{2}=6$ distinct pairs of
simultaneously active neurons, and if all five neurons were active we would
have a maximum of $\tbinom{5}{2}=10$ distinct pairs; the normalized number
of pairs is therefore $6/10$. Note that the first $m$ factorial moments
linearly determine the first $m$ moments and vice versa, so fixing the ones is
equivalent to fixing the others. Now assume that these $n$ neurons are a
sample from a full population of $N$ neurons, with possible normalized
activity $\yA \in\set{0,1/N,\dotsc,1}$. We have this extremely convenient
property: \emph{The first $n$ normalized factorial moments for the sample
  and for the full population are numerically identical}:
\begin{equation}
  \label{eq:equal_m_factorialmoments}
  \E\condt[\big]{\tbinom{n\ya}{m}\;\tbinom{n}{m}^{-1}}{\yH} =
  \E\condt[\big]{\tbinom{N\yA}{m}\;\tbinom{N}{m}^{-1}}{\yH},
  \qquad m \le n.
\end{equation}

We can hence apply the maximum-entropy method to obtain a distribution for
the full population by constraining its $m$th factorial moment to be equal
to the sample's recorded average of $\tbinom{n\ya}{m}\;\tbinom{n}{m}^{-1}$,
for as many $m$ as we please with $1\le m \le n$. If $\ya_{t}$ is the
normalized total activity of the sample during the time bin $t$, and the
recording with $T$ time bins, the constraint on the distribution
$\pf(\yA \| \yH)$ for the full population is
\begin{equation}
  \label{eq:final_constraint}
  \frac{1}{T}\sum_{t}\binom{n\ya_{t}}{m}\,\binom{n}{m}^{-1}
  =
   \sum_{\yA}
  \binom{N \yA}{m}\,  \binom{N}{m}^{-1}\, \pf(\yA \| \yH)
\end{equation}
for any $m$s we wish.
